Simulations of gravity-driven flow of binary liquids in microchannels

### 2011-03-26 Text/paper.tex
Working with corrections.

\section{Lattice Boltzmann binary liquid model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discrete velocity set.
The LBE is typically implemented as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{ collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
relaxation parameter, and $F_i$ is the external force population.{\color{red} The force population
is connected with the force and is implemented in the current work due to work \cite{guo}.}

@@

The model we use is a two-dimensional nine-velocity (D2Q9) model,
with equilibrium populations \cite{pooley-contact}:
\begin{equation}
\label{set:equilibrium:binary}
\begin{aligned}
&f_i^{eq}&&=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr)\\
&&&+k\bigl(w_i^{xx} (\partial_x \phi)^2+w_i^{yy} (\partial_y \phi)^2 +w_i^{xy} \partial_x
\phi \partial_y \phi \bigr), i=1\div8\\
&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div8 \\
&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
tension; $A$ is the parameter of the free-energy model. The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. 
Parameters specific to the D2Q9 grid are the weights
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$.  The weights related to the
inclusion of the surface tension coefficient into the equations are as follows:
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. The set of equations (\ref{set:equilibrium:binary}) restores the
macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}= F_{\alpha}
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields, {\color{red}
$P_{\alpha\beta}=\Bigl(p_0-k\phi \Delta \phi -\frac{k}{2}|\nabla \phi|^2\Bigr)\delta_{\alpha\beta}
+ k \partial_{\alpha} \phi \partial_{\beta} \phi$  \cite{pooley-contact}.}

The system allows the separation of the liquid
phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
relaxation time is taken as linearly dependent on the relaxation
times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
to change viscosity from the gas viscosity
$\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
accordingly.The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
A}{9}}$.

\section{Lattice Boltzmann benchmark}

@@

\includegraphics[width=0.97\textwidth]{Figures/benchmark.eps}

@@

\section{Results}
\subsection{The nondimensionalization and initialization procedure}
\label{section:init}
We now introduce the procedure of nondimensionalization, as well as
the initialization technique we used to determine the parameters necessary for
the simulation. The capillary number defined in Eq.~\ref{capillary:number:definition} governs the
interface thickness.
%The bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
%thickness is examined.
While parameters such as $\mu_{\mathrm{liq}}$, $\gamma$ and the capillary number $\Ca$ are given
or defined through the model parameters, the bubble velocity $U_{\mathrm{bubble}}$ is not given
explicitly.
In our simulations we approximate $U_{bubble}$ with the
Poiseuille profile center velocity, which is given as:
\begin{equation}
\begin{aligned}
U_{\mathrm{bubble}} \approx U_{\mathrm{liq}}=\frac{H_{\mathrm{eff}}^2}{8
\mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
U_{\mathrm{bubble}}\approx
\frac{{(N_y-2)}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
x } ,
\end{aligned}
\end{equation}
where $N_y$ is the grid number in the vertical direction. As will be discussed further, the
effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes to
mimic no-slip conditions at the walls.
For the Poiseuille profile one can estimate the body force (pressure gradient) through the
capillary number and grid resolution:
\begin{equation}
\label{poiseuille:velocity:center}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{(N_y-2)}^2}\gamma \Ca.
\end{aligned}
\end{equation}
In simulations the body force (\ref{poiseuille:velocity:center}) causes the bubble
velocity to differ from that implied by the Poiseuille relation because of the bubble presence in a
flow. That is
why an iteration procedure is needed to achieve the 
desired velocity and 
capillary number. 
However, the initialization body force given by (\ref{poiseuille:velocity:center}) is a good
starting approximation.

@@

{\color{red}
Note that the initialization procedure is a good guide of designing the simulations. For example
for the grid refinement, see Section \ref{section:grid:refinement}, we keep the same velocity
$U_{\mathrm{bubble}}$ and $\Ca$. This implies the conservation of the quantity
$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
x } = \mathrm{const}$. For the designing capillary range simulations, we use a proportionality law
to initialize body forces knowing results of just one simulation:
\begin{equation}
\begin{aligned}
&\Ca_{\mathrm{lit}} \propto U_{\mathrm{bubble}}\\
&U_{\mathrm{bubble}} \propto \frac{\mathrm{d}P}{\mathrm{d}x} H_{\mathrm{eff}}^2\\
&\Ca_{\mathrm{lit}} \propto \frac{\mathrm{d}P}{\mathrm{d} x} H_{\mathrm{eff}}^2 \text{ or }\\
&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{\Ca_{\mathrm{lit}}}{H_{\mathrm{eff}}^2},
\end{aligned}
\end{equation}
where the subscript ,,lit'' stands for the predicted capillary number
\cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
involved because all simulations are conducted on the same grid.
}
\subsection{Grid refinement}
\label{section:grid:refinement}
To properly estimate the interface resolution one needs to study the convergence as a function of
the grid resolution. To do that the grid resolution is varied while all remaining parameters,
including the bubble velocity and the capillary number, are fixed.  Our goal is to determine the 
ratio of the interface thickness to the 
film thickness at which results are no longer dependent on the grid resolution.

@@

{\color{red} After choosing the reference parameters, the grid refinement procedure
needs to keep the macroscopic parameters constant. One can refer to Section \ref{section:init}.}  
%It is easy to check
%that it yields the following quantity to
%be independent of grid size
%$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
%x } = \mathrm{const}$. 
We performed a number of simulations for the prescribed grids. The
simulation results in terms of grid numbers $N_x$, $N_y$, the film thickness $\delta$, center bubble
velocity $U_{bubble}$ are summarized in Table \ref{table:parameters:grid:refinement}.
The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
see that results converge for $H_{\mathrm{eff}}\geq 175$, and that with proper
initialization techniques, large enough time and different wall
wettabilities, results are different only in the 3rd digit even for
underresolved film thicknesses. To calculate how well the interface is
resolved, the ratio of the interface thickness to the film thickness is calculated. The
interface itself occupies approximately $5 \xi$, where
$\xi=\sqrt{k/A}=1$. The ratio of the interface thickness to the film thickness
$5\xi/H_{\mathrm{film}}$
 is shown in Table \ref{table:parameters:grid:refinement}. Based on these results one can conclude that
 the interface needs to be resolved as $40-50$ percent of the
expected film thickness for simulations to be grid independent. We further examine the
velocities in the center of the bubble to calculate the capillary number. One can see from Table
\ref{table:parameters:grid:refinement} that the bubble velocities are consistent and the
calculated capillary number corresponding to these velocities is $0.047$, which is {\color{red}
close to} the capillary number we aimed for. The corresponding difference can be
attributed to the Poiseuille profile body force initialization.% and the
 %density and viscosity ratios, which are much larger in the
%Bretherton problem.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Grid/norm_grid_profs.eps}
\caption{Grid-refined profiles for the effective
channel widths
$H_{\mathrm{eff}}=100,125,150,175,200,225$.$\delta$ is scaled on $H_{\mathrm{eff}}$ and $\delta=0$
corresponds to the wall location. The profiles were taken at $x=14$ (nondimensional coordinates).
Capillary number is $\Ca=0.05$. The capillary number obtained from simulations $\Ca=0.047$. 
%One can seeconvergent results even for the underresolved film
%thickness.
\label{fig:grid:profiles}}
\end{figure}
%Here will be a consolidated table
\begin{table}
\begin{tabularx}{0.97\textwidth}{|X|X|X|X|X|X|X|}
\hline
$N_x$&$N_y$&$\delta$&$U_\mathrm{bubble}$&$\frac{5\xi}{H_{\mathrm{film}}}$&$N_\mathrm{iter}$&{
\color{red} $Re$}\\
\hline
$1500$&$102$&$0.0694$&$0.0028$&$0.824$&$200000$&$0.44$\\
\hline
$1875$&$127$&$0.0690$&$0.0028$&$0.646$&$250000$&$0.53$\\
\hline
$2250$&$152$&$0.0675$&$0.0027$&$0.539$&$300000$&$0.63$\\
\hline
$2625$&$177$&$0.0674$&$0.0027$&$0.453$&$350000$&$0.73$\\
\hline
$3000$&$202$&$0.0662$&$0.0027$&$0.400$&$400000$&$0.83$\\
\hline
$3375$&$227$&$0.0664$&$0.0026$&$0.355$&$450000$&$0.91$\\
\hline
\end{tabularx}
\caption{The parameters and results for grid resolution. The simulated domain is
of size $N_x \times N_y$. $U_\mathrm{bubble}$ is the velocity of the interface at the fron tip of
the bubble.  $5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta (N_y-2)$ is the size of
the film in
lattice Boltzmann units. $Ca=0.05$ was taken to initialize the simulation.
\label{table:parameters:grid:refinement}}
\end{table}

@@

Another question that arises is the influence of the wall gradient on the bubble velocity and the
corresponding capillary number. The shear stress controlled by the wall phase gradient changes the
effective viscosity near the wall and the bubble velocity respectively. If the viscosity ratio is
high enough then the free surface motion of the bubble should not be dependent on the shear stress
near the wall.
One can see that if the film thickness is resolved properly the influence of the wall gradient can
be neglected. The velocities of the bubble center for profiles along $x$ direction are
presented in Table
\ref{table:parameters:wall:gradient}. 
%Fig. \ref{fig:velocities:wall:gradient} shows the center bubble velocity for different
%gradients. 
The average relative error for the mean velocity is $0.1\%$.
% Uncomment if you need this picture
% \begin{figure}
% \includegraphics[width=0.97\textwidth]{Figures/Wall/velocities_grad_profiles.eps}
% \caption{The center bubble velocity versus the phase gradient. One can neglect the influence of
% the wall once the film is properly resolved.
% \label{fig:velocities:wall:gradient}
% }
% \end{figure}
\begin{table}
\begin{tabularx}{\textwidth}{|X|X X X X X X X X X X X|}
\hline
$\scriptstyle \partial_n \phi$& $\scriptstyle -1.0$& $\scriptstyle -0.8$&
$\scriptstyle -0.6$&$\scriptstyle -0.4$&$\scriptstyle -0.2$&$\scriptstyle
0.0$&$\scriptstyle 0.2$&$\scriptstyle 0.4$&$\scriptstyle 0.6$&$\scriptstyle 0.8$&$\scriptstyle
1.0$\\
\hline
$\scriptstyle \delta$& $\scriptstyle 0.0633$& $\scriptstyle 0.0634$& $\scriptstyle 0.0634$&
$\scriptstyle 0.0634$& $\scriptstyle 0.0634$& $\scriptstyle 0.0634$& $\scriptstyle 0.0633$&
$\scriptstyle 0.0632$& $\scriptstyle 0.0631$ &$\scriptstyle \mathrm{N/A}$&$\scriptstyle
\mathrm{N/A}$\\
\hline
$\scriptstyle U_{bubble}$ &$\scriptstyle 0.0041$& $\scriptstyle 0.0042$& $\scriptstyle 0.0042$
&$\scriptstyle 0.0042$ & $\scriptstyle 0.0041$& $\scriptstyle 0.0041$ & $\scriptstyle 0.0042$ &
$\scriptstyle 0.0042$ & $\scriptstyle 0.0042$ & $\scriptstyle \mathrm{N/A}$ &$\scriptstyle
\mathrm{N/A}$\\
\hline
\end{tabularx}
\caption{The parameters and results for wall gradient effects on the film thickness. $U_{bubble}$
stands for the
center bubble velocity. $\delta$ is the interface thickness. $\partial_n \phi$ is the phase
gradient near the wall responsible for hydrophilic and hydrophobic behavior. The results are
calculated at the middle of the bubble after $200 000$ time iterations. $\Ca=0.05$ was taken to
initialize the simulations.
\label{table:parameters:wall:gradient}}
\end{table}

>>>>

\section{Lattice Boltzmann binary liquid model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discrete velocity set.
The LBE is typically implemented as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{ collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $f_i$ is the probability distribution function in the direction $\bm{c_i}$,
{\color{red} $f_i^{q}$ is the equilibrium probability distribution function}, $\omega$ is the
relaxation parameter, and $F_i$ is the external force population.{\color{red} The force population
is connected with the force and is implemented in the current work due to work \cite{guo}.}

@@

The model we use is a two-dimensional nine-velocity (D2Q9) model,
with equilibrium populations \cite{pooley-contact}:
\begin{equation}
\label{set:equilibrium:binary}
\begin{aligned}
&f_i^{eq}&&=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr)\\
&&&+k\bigl(w_i^{xx} (\partial_x \phi)^2+w_i^{yy} (\partial_y \phi)^2 +w_i^{xy} \partial_x
\phi \partial_y \phi \bigr), 1\leq i \leq 8\\
&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), 1\leq i \leq 8 \\
&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
tension; $A$ is the parameter of the free-energy model. The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$ {\color{red} with
the sound speed $c_s^2=1/3$}. 
Parameters specific to the D2Q9 grid are the weights
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$.  
{\color{red}
Other weights 
%related to the
%inclusion of the surface tension coefficient into the equations 
are as follows:}
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. The set of equations (\ref{set:equilibrium:binary}) restores the
macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}= F_{\alpha}
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields, {\color{red}
$P_{\alpha\beta}=\Bigl(p_0-k\phi \Delta \phi -\frac{k}{2}|\nabla \phi|^2\Bigr)\delta_{\alpha\beta}
+ k \partial_{\alpha} \phi \partial_{\beta} \phi$  \cite{pooley-contact}.The interface tension value
in the framework of the binary liquid model is $\gamma=\sqrt{\frac{8 k
A}{9}}$. The inclusion of the interface tension to the momentum flux tensor is done through the
coefficients $k$, $A$ and weights $w_i^{\alpha\beta}$.

Note that the first equation of system \ref{standard:implementation} simulates the continuity and
the Navier-Stokes equations, first two equations in (\ref{binary:fluid:system}). The second equation
of system \ref{standard:implementation} simulates the phase governing equation, third equation in
(\ref{binary:fluid:system}).} The system (\ref{binary:fluid:system}) allows the separation of the
liquid
phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
relaxation time is taken as linearly dependent on the relaxation
times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
to change viscosity from the gas viscosity
$\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
accordingly.

{\color{red}
While the lattice Boltzmann system has parameters as gas and liquid
viscosities, the surface tension and etc., those parameters are not the representative and
proportional quantities of the parameters in a physical system. The parameters of the lattice
Boltzmann are connected with the physical world's parameters only through the non-dimensional
numbers governing a physics of the problem. In this case it's a capillary number $\Ca$, which is
obtained from the physical world and then it is matched through the lattice Boltzmann quantities.
In fact, one should distinguish the lattice Boltzmann system and the physical system. The set of
the fluid and phase equations \eqref{binary:fluid:system} is valid in the lattice Boltzmann space
and in the physical domain. Therefore, one can substitute any quantity, i.e.
$U_{\mathrm{bubble}}$, in the physical units or in the lattice Boltzmann units as soon as the
capillary number is matched in the both of worlds.}

%\section{Lattice Boltzmann benchmark}
\section{Approach to benchmark}

@@

\includegraphics[width=0.97\textwidth]{Figures/benchmark_new.eps}

@@

\section{Results}
\subsection{The nondimensionalization and initialization procedure}
\label{section:init}
We now introduce the procedure of nondimensionalization, as well as
the initialization technique we used to determine the parameters necessary for
the simulation. The capillary number defined in Eq.~\ref{capillary:number:definition} governs the
interface thickness.
%The bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
%thickness is examined.
While parameters such as $\mu_{\mathrm{liq}}$, $\gamma$ and the capillary number $\Ca$ are given
or defined through the model parameters, the bubble velocity $U_{\mathrm{bubble}}$ is not given
explicitly.
In our simulations we approximate $U_{bubble}$ with the
Poiseuille profile center velocity, which is given as:
\begin{equation}
\begin{aligned}
U_{\mathrm{bubble}} \approx \frac{H_{\mathrm{eff}}^2}{8
\mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
U_{\mathrm{bubble}}\approx
\frac{{(N_y-2)}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
x } ,
\end{aligned}
\end{equation}
where $N_y$ is the grid number in the vertical direction. As will be discussed further, the
effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes to
mimic no-slip conditions at the walls.
For the Poiseuille profile one can estimate the body force (pressure gradient) through the
capillary number and grid resolution:
\begin{equation}
\label{poiseuille:velocity:center}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{(N_y-2)}^2}\gamma \Ca.
\end{aligned}
\end{equation}
In simulations the body force (\ref{poiseuille:velocity:center}) causes the bubble
velocity to differ from that implied by the Poiseuille relation because of the bubble presence in a
flow. That is
why an iteration procedure is needed to achieve the 
desired velocity and 
capillary number. 
However, the initialization body force given by (\ref{poiseuille:velocity:center}) is a good
starting approximation.

@@

{\color{red}
Note that the initialization procedure is a good guide of designing the simulations. For example
for the grid refinement, Section \ref{section:grid:refinement}, we keep the same velocity
$U_{\mathrm{bubble}}$ and $\Ca$. This implies the conservation of the quantity
$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
x } = \mathrm{const}$. For the designing capillary range simulations, Section
\ref{section:capillary:range}, we use a proportionality law to initialize body forces knowing
results of just one simulation:
\begin{equation}
\begin{aligned}
&\Ca_{\mathrm{lit}} \propto U_{\mathrm{bubble}}\\
&U_{\mathrm{bubble}} \propto \frac{\mathrm{d}P}{\mathrm{d}x} H_{\mathrm{eff}}^2\\
&\Ca_{\mathrm{lit}} \propto \frac{\mathrm{d}P}{\mathrm{d} x} H_{\mathrm{eff}}^2 \text{ or }\\
&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{\Ca_{\mathrm{lit}}}{H_{\mathrm{eff}}^2},
\end{aligned}
\end{equation}
where the subscript ,,lit'' stands for the predicted capillary number
\cite{giavedoni-numerical,heil-bretherton}. 
}
\subsection{Grid refinement}
\label{section:grid:refinement}
To properly estimate the interface resolution one needs to study the convergence as a function of
the grid resolution. To do that the grid resolution is varied while all remaining parameters,
including the bubble velocity and the capillary number, are fixed.  Our goal is to determine the 
ratio of the interface thickness to the 
film thickness at which results are no longer dependent on the grid resolution.

@@

{\color{red} After choosing the reference parameters, the grid refinement procedure
needs to keep the macroscopic parameters constant. One can refer to Section \ref{section:init}.}  
%It is easy to check
%that it yields the following quantity to
%be independent of grid size
%$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
%x } = \mathrm{const}$. 
We performed a number of simulations for the prescribed grids. The
simulation results in terms of grid numbers $N_x$, $N_y$, the film thickness $\delta$, center bubble
velocity $U_{bubble}$ are summarized in Table \ref{table:parameters:grid:refinement}.
The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
see that results converge for $H_{\mathrm{eff}}\geq 175$, and that with proper
initialization techniques, large enough time and different wall
wettabilities, results are different only in the 3rd digit even for
underresolved film thicknesses. To calculate how well the interface is
resolved, the ratio of the interface thickness to the film thickness is calculated. The
interface itself occupies approximately $5 \xi$, where
$\xi=\sqrt{k/A}=1$. The ratio of the interface thickness to the film thickness
$5\xi/H_{\mathrm{film}}$
 is shown in Table \ref{table:parameters:grid:refinement}. Based on these results one can conclude that
 the interface needs to be resolved as $40-50$ percent of the
expected film thickness for simulations to be grid independent. We further examine the
velocities in the center of the bubble to calculate the capillary number. One can see from Table
\ref{table:parameters:grid:refinement} that the bubble velocities are consistent and the
calculated capillary number corresponding to these velocities is $0.047$, which is {\color{red}
close to} the capillary number we aimed for. The corresponding difference can be
attributed to the Poiseuille profile body force initialization.% and the
 %density and viscosity ratios, which are much larger in the
%Bretherton problem.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Grid/norm_grid_profs.eps}
\caption{Grid-refined profiles for the effective
channel widths
$H_{\mathrm{eff}}=100,125,150,175,200,225$.$\delta$ is scaled on $H_{\mathrm{eff}}$ and $\delta=0$
corresponds to the wall location. The profiles were taken at $x=14$ (nondimensional coordinates).
Capillary number is $\Ca=0.05$. The capillary number obtained from simulations $\Ca=0.047$. 
%One can seeconvergent results even for the underresolved film
%thickness.
\label{fig:grid:profiles}}
\end{figure}
%Here will be a consolidated table
\begin{table}
\begin{tabularx}{0.97\textwidth}{|X|X|X|X|X|X|X|}
\hline
$N_x$&$N_y$&$\delta$&$U_\mathrm{bubble}$&$\frac{5\xi}{H_{\mathrm{film}}}$&$N_\mathrm{iter}$&{
\color{red} $Re$}\\
\hline
$1500$&$102$&$0.0694$&$0.0028$&$0.824$&$200000$&$0.44$\\
\hline
$1875$&$127$&$0.0690$&$0.0028$&$0.646$&$250000$&$0.53$\\
\hline
$2250$&$152$&$0.0675$&$0.0027$&$0.539$&$300000$&$0.63$\\
\hline
$2625$&$177$&$0.0674$&$0.0027$&$0.453$&$350000$&$0.73$\\
\hline
$3000$&$202$&$0.0662$&$0.0027$&$0.400$&$400000$&$0.83$\\
\hline
$3375$&$227$&$0.0664$&$0.0026$&$0.355$&$450000$&$0.91$\\
\hline
\end{tabularx}
\caption{The parameters and results for grid resolution. The simulated domain is
of size $N_x \times N_y$. $U_\mathrm{bubble}$ is the velocity of the interface at the fron tip of
the bubble.  $5\xi$ is the interface thickness. $H_{\mathrm{film}}=\delta (N_y-2)$ is the size of
the film in
lattice Boltzmann units. $\Ca=0.05$ was taken to initialize the simulation.
\label{table:parameters:grid:refinement}}
\end{table}

@@

Another question that arises is the influence of the wall gradient on the bubble velocity and the
corresponding capillary number. The shear stress controlled by the wall phase gradient changes the
effective viscosity near the wall and the bubble velocity respectively. If the viscosity ratio is
high enough then the free surface motion of the bubble should not be dependent on the shear stress
near the wall.
One can see that if the film thickness is resolved properly the influence of the wall gradient can
be neglected. The velocities of the bubble center for profiles along $x$ direction are
presented in Table
\ref{table:parameters:wall:gradient}. 
%Fig. \ref{fig:velocities:wall:gradient} shows the center bubble velocity for different
%gradients. 
The average relative error for the mean velocity is $0.1\%$.
% Uncomment if you need this picture
% \begin{figure}
% \includegraphics[width=0.97\textwidth]{Figures/Wall/velocities_grad_profiles.eps}
% \caption{The center bubble velocity versus the phase gradient. One can neglect the influence of
% the wall once the film is properly resolved.
% \label{fig:velocities:wall:gradient}
% }
% \end{figure}
\begin{table}
\begin{tabularx}{\textwidth}{|X|X X X X X X X X X X X|}
\hline
$\scriptstyle \partial_n \phi$& $\scriptstyle -1.0$& $\scriptstyle -0.8$&
$\scriptstyle -0.6$&$\scriptstyle -0.4$&$\scriptstyle -0.2$&$\scriptstyle
0.0$&$\scriptstyle 0.2$&$\scriptstyle 0.4$&$\scriptstyle 0.6$&$\scriptstyle 0.8$&$\scriptstyle
1.0$\\
\hline
$\scriptstyle \delta$& $\scriptstyle 0.0633$& $\scriptstyle 0.0634$& $\scriptstyle 0.0634$&
$\scriptstyle 0.0634$& $\scriptstyle 0.0634$& $\scriptstyle 0.0634$& $\scriptstyle 0.0633$&
$\scriptstyle 0.0632$& $\scriptstyle 0.0631$ &$\scriptstyle \mathrm{N/A}$&$\scriptstyle
\mathrm{N/A}$\\
\hline
$\scriptstyle U_{bubble}$ &$\scriptstyle 0.0041$& $\scriptstyle 0.0042$& $\scriptstyle 0.0042$
&$\scriptstyle 0.0042$ & $\scriptstyle 0.0041$& $\scriptstyle 0.0041$ & $\scriptstyle 0.0042$ &
$\scriptstyle 0.0042$ & $\scriptstyle 0.0042$ & $\scriptstyle \mathrm{N/A}$ &$\scriptstyle
\mathrm{N/A}$\\
\hline
\end{tabularx}
\caption{The parameters and results for wall gradient effects on the film thickness.
$U_{\mathrm{bubble}}$
stands for the interface velocity. $\delta$ is the film thickness. $\partial_n \phi$ is the phase
gradient near the wall responsible for hydrophilic and hydrophobic behavior. The results are
calculated at the middle of the bubble after $200 000$ time iterations. $\Ca=0.05$ was taken to
initialize the simulations.
\label{table:parameters:wall:gradient}}
\end{table}
